---
title: ReadMe package <br> for Content Analysis
author: "Ginevra Carbone, Domagoj Korais"
date: "28 Giugno 2018"
output:
  ioslides_presentation:
    fig_width: 7
    incremental: yes
    smaller: yes
    widescreen: yes
  html_document:
    toc: yes
  slidy_presentation:
    incremental: yes
editor_options:
  chunk_output_type: inline
subtitle: A theoretical and practical evaluation
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(cache = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)

set.seed(123)
mainwd <- getwd()
# setwd("/home/ginevracoal/MEGA/Università/DSSC/semester_2/statistical_methods_for_data_science/statMethodsProject/source")
```

```{r libraries, include=FALSE}
library(dplyr)
library(ggplot2)
library(ReadMe)
library(gridExtra)
```

## Introduction | Content Analysis {.build}

<div class="centered">
What is **Sentiment Analysis**? 
</div>
It is the process of *categorizing* opinions about a specific sentence or text. 

<br>

<div class="columns-2">

<div align = "center">
**Classification methods**
</div>

*Goal:* increasing the accuracy of individual classifications,
which is the highest number of correct predictions.

Suppose that missclassifications mostly occur in some specific categories.
This is the case of a *statistical bias*: average difference between true and estimated proportion of documents in each category.

<!-- <br> -->
<div align="center">
Here comes **Content Analysis**
</div>

*Goal*: getting approximately unbiased estimates of category proportions.

Notice that there exists *no necessary connection* between low misclassification rates and low bias in each category, except at the extremes of classification.

## Introduction | Amazon reviews dataset {.build}

An example of categorical proportions on *Amazon reviews*.

```{r}
cds_prop <- readRDS("source/results/cds_prop.rds")
instr_prop <- readRDS("source/results/instr_prop.rds")
cell_prop <- readRDS("source/results/cell_prop.rds")
elec_prop <- readRDS("source/results/elec_prop.rds")
```

<div class="centered">
```{r}
cds_plot <- cds_prop %>% ggplot(aes(x = year, y = prop, group = factor(rating), color = rating)) + 
  geom_point() + geom_line() +
  scale_color_brewer(palette="Set2")+
  xlab("review time") + ylab("proportion of reviews")+
  labs(title="CDs and Vinyls ratings")+
  theme_minimal()

instr_plot <- instr_prop %>% 
  ggplot(aes(x = year, y = prop, group = factor(rating), color = rating)) + 
  geom_point() + geom_line() +
  scale_color_brewer(palette="Set2")+
  xlab("review time") + ylab("proportion of reviews")+
  labs(title="Musical instruments ratings")+
    theme_minimal()

cell_plot <- cell_prop %>% 
  ggplot(aes(x = year, y = prop, group = factor(rating), color = rating)) + 
  geom_point() + geom_line() +
  scale_color_brewer(palette="Set2")+
  xlab("review time") + ylab("proportion of reviews")+
  labs(title="Cellphones ratings")+
  theme_minimal()

elec_plot <- elec_prop %>% 
  ggplot(aes(x = year, y = prop, group = factor(rating), color = rating)) + 
  geom_point() + geom_line() + 
  scale_color_brewer(palette="Set2")+
  theme_minimal()+
  xlab("review time") + ylab("proportion of reviews")+
  labs(title="Electronics ratings")

grid.arrange(cds_plot,instr_plot, cell_plot, elec_plot, ncol=2, nrow=2)
```

## Theoretical background | Analyzing text statistically {.build}

Documents need to be **preprocessed** by removing lowercase, punctuation, spaces, stop words, applying stemming and creating the so called "bag of words".

<br>

We will use the following *notation*.

- $j=1,\dots,J$ are document categories;
- there are two sets of text documents:
    - labeled training set $i=1,\dots,n$;
    - unlabeled test set $l=1,\dots,L$;
- $D_i = j$ if document $i$ belongs to category $j$;
- $\{S_{i1},\dots,S_{iK}\}$ is the set of stem words for all documents, where $S_{ik}=\delta_{ik}$.

## Theoretical background | Analyzing text statistically {.build}

<div align="center">
**Individual classifications**

$\{D_1,\dots,D_L\}$

vs

**Proportions of classifications** 

$P(D) = \{P(D=1),\dots,P(D=J)\}$
</div>

<br>

where $P(D=j) = \frac{1}{L} \sum_{l=1}^{L} I(D_l = j)$ is the proportion of documents belonging to category $j$.


## Theoretical background | The method {.build}

The method provided by `ReadMe` package allows to calculate these estimates **without performing an individual classification**, let's understand how.

Let $\hat{D}$ be the estimated classification for the documents. Then the *law of total probability* allows to express the proportion of documents estimated to be in category $j$ in terms of the **missclassification probabilities** $P(\hat{D}=j \,|\, D=j')$ and the **true proportion** as
$$
P(\hat{D}=j)=\sum_{j'=1}^{L} P(\hat{D}=j \,|\, D=j') P(D=j').
$$


The idea is to avoid computing $P(\hat{D}=j)$, based on the fact that each text is written by a person with a specific *opinion*.


## Theoretical background | The method {.build}

Since the words chosen are by definition a function of the document category and the categories partition our sample space, the probability of each of the $2^k$ stem word combinations (*stem profiles*) occuring is
$$
P(S=s)=\sum_{j=1}^{L} P(S=s \,|\, D=j) P(D=j)
$$
where $P(S=s | D=j)$ is the probability of $s$ occuring within a document of category $D$.

We get the equivalent matrix expression  
$$
\underset{2^K\times1}{P(S)} = \underset{2^K\times J}{P(S\,|\,D)} \underset{J\times1}{P(D)}.
$$

## Theoretical background | The method {.build}

The **fundamental assumption** for this model is the proportion $P(S=s \,|\, D=j)$ being the same for both training and test set, so it assumes that the documents in the labeled set contain a sufficient number of examples to describe the language used in each category.

We can now directly solve this equation using regression on the coefficients $P(D)$, but there are two relevant problems we have to pay attention to:

- $K$ can be very large, so $2^K$ makes this model **computationally expensive**;
- The matrix $P(S=s \,|\, D=j)$ is **sparse**, since $n$ is usually much smaller than $2^K$.

<br>
<div align="center">
How do we avoid these problems?
</div>

We randomly choose a subset of words, as seen in [2].

## Inside ReadMe package | functions {.build}

The package contains three main functions:

- `undergrad()` translates a set of texts stored in a single folder into a dataset, where each text is a row and each word is a column, the so called "**document term matrix**", based on a specific threshold;

- `preprocess()` removes all columns with null variance, i.e. extremely **rare words**;

- `readme()` computes the **real proportions** on the train set and the **estimated proportions** on the test set. It optionally computes **bootstrap intervals** too.

Let's take a better look into these function through an example.

## Inside ReadMe package | Rotten Tomatoes dataset {.build}

The first dataset we analyzed contains $8529$ movie reviews from the famous website *www.rottentomatoes.com*, labeled with ratings from $1$ to $5$.

The inputs for `undergrad()` are the text files containing all the reviews and a data frame containing file names, classification labels, and a (randomly generated) binary value separating train set from test set.

```{r}
head(readRDS("source/results/rotten_control.rds"))
```

`preprocess()` found no invariant columns, probably due to the large dimension of our dataset.

## Inside ReadMe package | Rotten Tomatoes dataset {.build}

After applying the main function `readme()`, we computed the same proportions from the results of two diffent classification methods: support vector machines using **linear** and **radial** kernels respectively.

<div align="center">
```{r fig.height=4.2}
rotten <- readRDS("source/results/rotten.rds")

ggplot(rotten, aes(true, est, color = rating, shape = type)) +
  geom_point(size = 2) +
  xlab(expression(P(D))) +
  ylab(paste("Estimated ", expression(P(D)))) +
  geom_abline(slope = 1, intercept = 0) +
  xlim(c(0,0.6)) + ylim(c(0,0.6)) +
  scale_color_brewer(palette="Set2")+
  scale_shape_manual(values = c(0, 2, 16))+
  theme_minimal()
```
</div>

## Inside ReadMe package | Rotten Tomatoes dataset {.build}

The following tables compares the **mean absolute errors** on the proportions for these methods.

```{r}
readRDS("source/results/mae_table.rds")
```

## Doma

## Bibliography

[1] "A Method of Automated Nonparametric Content Analysis for Social Science", D. J. Hopkins, G. King, 2010.
[2] "Verbal Autopsy Methods with Multiple Causes of Death", King, Gary and Ying Lu, 2008.