---
title: ReadMe package <br> for Content Analysis
author: "Ginevra Carbone, Domagoj Korais"
date: "28 Giugno 2018"
output:
  ioslides_presentation:
    fig_width: 7
    incremental: yes
    smaller: yes
    widescreen: yes
  html_document:
    toc: yes
  slidy_presentation:
    incremental: yes
editor_options:
  chunk_output_type: inline
subtitle: A theoretical and practical evaluation
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(cache = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)

set.seed(123)
mainwd <- getwd()
# setwd("/home/ginevracoal/MEGA/Università/DSSC/semester_2/statistical_methods_for_data_science/statMethodsProject/source")
```

```{r libraries, include=FALSE}
library(dplyr)
library(ggplot2)
library(ReadMe)
library(gridExtra)
```

## Introduction | Content Analysis {.build}

<div class="centered">
What is **Sentiment Analysis**? 
</div>
It is the process of *categorizing* opinions about a specific sentence or text. 

<br>

<div class="columns-2">

<div align = "center">
**Classification methods**
</div>

*Goal:* increasing the accuracy of individual classifications,
which is the highest number of correct predictions.

Suppose that missclassifications mostly occur in some specific categories.
This is the case of a *statistical bias*: average difference between true and estimated proportion of documents in each category.

<!-- <br> -->
<div align="center">
Here comes **Content Analysis**
</div>

*Goal*: getting approximately unbiased estimates of category proportions.

Notice that there exists *no necessary connection* between low misclassification rates and low bias in each category, except at the extremes of classification.


## Theoretical background | Analyzing text statistically {.build}

Documents need to be **preprocessed** by removing lowercase, punctuation, spaces, stop words, applying stemming and creating the so called "bag of words".

<br>

We will use the following *notation*.

- $j=1,\dots,J$ are document categories;
- there are two sets of text documents:
    - labeled training set $i=1,\dots,n$;
    - unlabeled test set $l=1,\dots,L$;
- $D_i = j$ if document $i$ belongs to category $j$;
- $\{S_{i1},\dots,S_{iK}\}$ is the set of stem words for all documents, where $S_{ik}=\delta_{ik}$.

## Theoretical background | Analyzing text statistically {.build}

<div align="center">
**Individual classifications**

$\{D_1,\dots,D_L\}$

vs

**Proportions of classifications** 

$P(D) = \{P(D=1),\dots,P(D=J)\}$
</div>

<br>

where $P(D=j) = \frac{1}{L} \sum_{l=1}^{L} I(D_l = j)$ is the proportion of documents belonging to category $j$.


## Theoretical background | The method {.build}

The method provided by `ReadMe` package allows to calculate these estimates **without performing an individual classification**, let's understand how.

Let $\hat{D}$ be the estimated classification for the documents. Then the *law of total probability* allows to express the proportion of documents estimated to be in category $j$ in terms of the **missclassification probabilities** $P(\hat{D}=j \,|\, D=j')$ and the **true proportion** as
$$
P(\hat{D}=j)=\sum_{j'=1}^{L} P(\hat{D}=j \,|\, D=j') P(D=j').
$$


The idea is to avoid computing $P(\hat{D}=j)$, based on the fact that each text is written by a person with a specific *opinion*.


## Theoretical background | The method {.build}

Since the words chosen are by definition a function of the document category and the categories partition our sample space, the probability of each of the $2^k$ stem word combinations (*stem profiles*) occuring is
$$
P(S=s)=\sum_{j=1}^{L} P(S=s \,|\, D=j) P(D=j)
$$
where $P(S=s | D=j)$ is the probability of $s$ occuring within a document of category $D$.

We get the equivalent matrix expression  
$$
\underset{2^K\times1}{P(S)} = \underset{2^K\times J}{P(S\,|\,D)} \underset{J\times1}{P(D)}.
$$

## Theoretical background | The method {.build}

The **fundamental assumption** for this model is the proportion $P(S=s \,|\, D=j)$ being the same for both training and test set, so it assumes that the documents in the labeled set contain a sufficient number of examples to describe the language used in each category.

We can now directly solve this equation using regression on the coefficients $P(D)$, but there are two relevant problems we have to pay attention to:

- $K$ can be very large, so $2^K$ makes this model **computationally expensive**;
- The matrix $P(S=s \,|\, D=j)$ is **sparse**, since $n$ is usually much smaller than $2^K$.

<br>
<div align="center">
How do we avoid these problems?
</div>

We randomly choose a subset of words, as seen in [2].

## Inside ReadMe package | functions {.build}

The package contains three main functions:

- `undergrad()` translates a set of texts stored in a single folder into a dataset, where each text is a row and each word is a column, the so called "**document term matrix**", based on a specific threshold;

- `preprocess()` removes all columns with null variance, i.e. extremely **rare words**;

- `readme()` computes the **real proportions** on the train set and the **estimated proportions** on the test set. It optionally computes **bootstrap intervals** too.

Let's take a better look into these function through an example.

## Inside ReadMe package | Rotten Tomatoes dataset {.build}

The first dataset we analyzed contains $8529$ movie reviews from the famous website *www.rottentomatoes.com*, labeled with ratings from $1$ to $5$.

The inputs for `undergrad()` are the text files containing all the reviews and a data frame containing file names, classification labels, and a (randomly generated) binary value separating train set from test set.

```{r}
head(readRDS("source/results/rotten_control.rds"))
```

`preprocess()` found no invariant columns, probably due to the large dimension of our dataset.

## Inside ReadMe package | Rotten Tomatoes dataset {.build}

After applying the main function `readme()`, we computed the same proportions from the results of two diffent classification methods: support vector machines using **linear** and **radial** kernels respectively.

<div align="center">
```{r fig.height=4.2}
rotten <- readRDS("source/results/rotten.rds")

ggplot(rotten, aes(true, est, color = rating, shape = type)) +
  geom_point(size = 2) +
  xlab(expression(P(D))) +
  ylab(paste("Estimated ", expression(P(D)))) +
  geom_abline(slope = 1, intercept = 0) +
  xlim(c(0,0.6)) + ylim(c(0,0.6)) +
  scale_color_brewer(palette="Set2")+
  scale_shape_manual(values = c(0, 2, 16))+
  theme_minimal()
```
</div>


```{r import dataset}
#Importing data
library(tidyverse)
#   rename(rating=overall)
# saveRDS(reviews, "../dataset/amazon_CDs_and_vinyl.rds")
reviews=readRDS("source/results/amazon_CDs_and_vinyl.rds")
#contiamo le parole nei testi, dirty trick, conto gli spazi e aggiungo 1
reviews=reviews%>%
  mutate(number_words= (str_count(reviewText,pattern = " ")+1) )
#class distribution plot----
```


## Inside ReadMe package | Rotten Tomatoes dataset {.build}

The following tables compares the **mean absolute errors** on the proportions for these methods.
```{r}
readRDS("source/results/mae_table.rds")
```

## Testing Readme package | Amazon reviews dataset {.build}

Let's now take a look at some real examples of categorical proportions.

```{r}
cds_prop <- readRDS("source/results/cds_prop.rds")
instr_prop <- readRDS("source/results/instr_prop.rds")
cell_prop <- readRDS("source/results/cell_prop.rds")
elec_prop <- readRDS("source/results/elec_prop.rds")
```

<div class="centered">
```{r fig.width=10}
cds_plot <- cds_prop %>% ggplot(aes(x = year, y = prop, group = factor(rating), color = rating)) + 
  geom_point() + geom_line() +
  scale_color_brewer(palette="Set2")+
  xlab("review time") + ylab("proportion of reviews")+
  labs(title="CDs and Vinyls ratings")+
  theme_minimal()

instr_plot <- instr_prop %>% 
  ggplot(aes(x = year, y = prop, group = factor(rating), color = rating)) + 
  geom_point() + geom_line() +
  scale_color_brewer(palette="Set2")+
  xlab("review time") + ylab("proportion of reviews")+
  labs(title="Musical instruments ratings")+
    theme_minimal()

cell_plot <- cell_prop %>% 
  ggplot(aes(x = year, y = prop, group = factor(rating), color = rating)) + 
  geom_point() + geom_line() +
  scale_color_brewer(palette="Set2")+
  xlab("review time") + ylab("proportion of reviews")+
  labs(title="Cellphones ratings")+
  theme_minimal()

elec_plot <- elec_prop %>% 
  ggplot(aes(x = year, y = prop, group = factor(rating), color = rating)) + 
  geom_point() + geom_line() + 
  scale_color_brewer(palette="Set2")+
  theme_minimal()+
  xlab("review time") + ylab("proportion of reviews")+
  labs(title="Electronics ratings")

grid.arrange(cds_plot,instr_plot, cell_plot, elec_plot, ncol=2, nrow=2)
```

## Testing Readme package | An overview{.build}

<div class="centered">
Tested on  **Amazon CD and vinyls** dataset
</div>


<br>

<div class="columns-2">

<div align = "center">
**The dataset**
</div>

- Provided by University of California, San Diego

- $1097592$ reviews

- five different labels

- Timespan of various years

<!-- <br> -->
<div align="center">
Performed **tests**
</div>

- Influence of dimension of the training set

- Influence of words number in document

- Assessing optimal parameters


## Main characteristics of the dataset | Words and ratings distributions


<div align="center">
```{r Distribution number of words, fig.width=10}

#plot dei risultati
reviews%>%filter(number_words<=1000)%>%
ggplot(aes(number_words))+
  geom_histogram()+
  ggtitle(" Number of words count")+
   xlab("Number of words")+
  ylab("Total Count")+
  xlim(0, 1000)+
  theme_minimal() ->p1

#plot dei risultati
reviews%>%filter(number_words<=50)%>%
ggplot(aes(number_words))+
  geom_histogram()+
  ggtitle(" Number of words count")+
   xlab("Number of words")+
   ylab("")+
  xlim(0, 50)+
  theme_minimal()->p2

plot_distribution_ratings=function(reviews){
classes=reviews%>%
  select(rating)%>%
  group_by(rating)%>%
  summarise(total=n())%>%
  mutate(percentage=total/sum(total))

p=ggplot(classes,aes(rating,total,fill=as.factor(rating)))+
  geom_col()+
  labs(fill="Rating",
       title=" Ratings count")+
  xlab("Rating")+
  ylab("")+
  scale_color_brewer(palette="Set2")+
  theme_minimal()
return(p)
}

p3=plot_distribution_ratings(reviews)

grid.arrange(p1,p2,p3,nrow=1,ncol=3)
```
</div>



## Relationship between number of words and ratings

<div class="columns-2">
```{r number of words, fig.width=4.5}


#check correlation length-rating----
reviews%>%
  filter( number_words<=400)%>%
ggplot(aes(x=as.factor(rating),y=number_words))+
  geom_boxplot()+
  labs(title="Relationship between number of words and rating",
       subtitle=" Number of words < 400.",
       caption="Source: University of California, San Diego")+
  ylab("Number of words")+
  xlab("Rating")+
  theme_minimal()

```

<br>

- There is no significant relationship between the number of words in the text and the final rating;

- It's possible to use this dataset to study the *influence of the number of words* on it's performace.

</div>

## Testing Readme package | Influence of dimension of the training set 
<div align=center>
```{r amazon training size determination}
library(tidyverse)
#import dati calcolati altro pc con m=10 mm=10
super_avg=readRDS("source/results/risultati_run_amazon_3h_nonliperdere.rds")
m=10
mm=10
#valori medi e deviazione standard dalla lista
media_rmse=data.frame(media=1:m)
sd_rmse = data.frame(sd=1:m)
training = data.frame(training = 1:m)
for (i in 1:m){
media_rmse[i,]=mean(super_avg[[i]]$avg)
sd_rmse[i,]=sd(super_avg[[i]]$avg)/mm^(1/2)
training[i,]=round(mean(super_avg[[i]]$training))
}
results=data.frame(training=training,media=media_rmse,sd=sd_rmse)


##plotting results
ggplot(results,aes(training,media))+
  geom_linerange(aes(ymin=media-sd, ymax=media+sd)) +
  geom_line()+
  labs(
  title="Avg. RMSE by elements in training set",
  subtitle=paste("Total number of texts:10000,10 run for each training value" ),
  caption=" Subset with 2000 elements for each class")+
  ylab("Mean")+
  scale_color_brewer(palette="Set2")+
  theme_minimal()
  
```

</div>

## Testing readme package | Influence of number of words per document
<div class="columns-2">



```{r frasi extra corte0, fig.width=4.5}
results=readRDS("source/results/results_extra_corte_1_15_cd.rds")
    ################plotting mod no bootstrap

    valori_fit = results$est.CSMF
    valori_veri = results$true.CSMF
    #sd_fit = results$CSMF.se
    #converto in dataframe
    valori_fit = as_tibble(as.list(valori_fit))
    valori_veri = as_tibble(as.list(valori_veri))
    #sd_fit = as_tibble(as.list(sd_fit))
    valori_plot=gather(valori_fit,type)
    valori_plot_veri=gather(valori_veri,type)
    #valori_sd=gather(sd_fit,type)
    #typeof(results$est.CSMF)
   
    #valori_sd
    valori_tot=inner_join(valori_plot,valori_plot_veri,by="type",suffix=c(".est",".real"))
    #valori_tot=inner_join(valori_tot,valori_sd)
    #valori_tot
valori_tot=rename(valori_tot,rating=type)
    p=ggplot(data=valori_tot,aes(x=value.real,y=value.est,label = value.est))+
      geom_point(aes(color=rating))+
    #  geom_linerange(aes(ymin=value.est-value, ymax=value.est+value,color=type)) +
     # ggtitle(" Computed vs fitted values for Amazon music dataset, frasi corte")+
      ggplot2::labs(
        title = "Number of words <=15",
        subtitle = paste("total number of elements:",9898,",","of which training:",1000))+
      # geom_smooth(method=lm)+
      xlab("Real value")+
      ylab("Computed value")+
      theme_minimal()+
      geom_abline(slope = 1,intercept = 0)

    plot(p)
    ############################
```


<br>

```{r frasi  corte0, fig.width=4.5}
results=readRDS("source/results/results_corte_16_50_cd.rds")
    ################plotting mod no bootstrap

    valori_fit = results$est.CSMF
    valori_veri = results$true.CSMF
    #sd_fit = results$CSMF.se
    #converto in dataframe
    valori_fit = as_tibble(as.list(valori_fit))
    valori_veri = as_tibble(as.list(valori_veri))
    #sd_fit = as_tibble(as.list(sd_fit))
    valori_plot=gather(valori_fit,type)
    valori_plot_veri=gather(valori_veri,type)
    #valori_sd=gather(sd_fit,type)
    #typeof(results$est.CSMF)
    
    #valori_sd
    valori_tot=inner_join(valori_plot,valori_plot_veri,by="type",suffix=c(".est",".real"))
    #valori_tot=inner_join(valori_tot,valori_sd)
    #valori_tot
valori_tot=rename(valori_tot,rating=type)
    p=ggplot(data=valori_tot,aes(x=value.real,y=value.est,label = value.est))+
      geom_point(aes(color=rating))+
    #  geom_linerange(aes(ymin=value.est-value, ymax=value.est+value,color=type)) +
     # ggtitle(" Computed vs fitted values for Amazon music dataset, frasi corte")+
      ggplot2::labs(
        title = "Number of words <=50 and >=16",
        subtitle = paste("total number of elements:",9898,",","of which training:",1000))+
      # geom_smooth(method=lm)+
      xlab("Real value")+
      ylab("Computed value")+
      theme_minimal()+
      geom_abline(slope = 1,intercept = 0)

    plot(p)
    ############################
```

</div>




## Testing readme package | Influence of number of words per document{.build}

<div class="columns-2">

```{r frasi lunghe, fig.width=4.5}
results=readRDS("source/results/results_lunghe_51_200_cd.rds")
    ################plotting mod no bootstrap

    valori_fit = results$est.CSMF
    valori_veri = results$true.CSMF
    #sd_fit = results$CSMF.se
    #converto in dataframe
    valori_fit = as_tibble(as.list(valori_fit))
    valori_veri = as_tibble(as.list(valori_veri))
    #sd_fit = as_tibble(as.list(sd_fit))
    valori_plot=gather(valori_fit,type)
    valori_plot_veri=gather(valori_veri,type)
    #valori_sd=gather(sd_fit,type)
   
    valori_tot=inner_join(valori_plot,valori_plot_veri,by="type",suffix=c(".est",".real"))
    #valori_tot=inner_join(valori_tot,valori_sd)
    #valori_tot
valori_tot=rename(valori_tot,rating=type)
    p=ggplot(data=valori_tot,aes(x=value.real,y=value.est,label = value.est))+
      geom_point(aes(color=rating))+
    #  geom_linerange(aes(ymin=value.est-value, ymax=value.est+value,color=type)) +
     # ggtitle(" Computed vs fitted values for Amazon music dataset, frasi corte")+
      ggplot2::labs(
          title = "Number of words <=200 and >=51",
        subtitle = paste("total number of elements:",10000,",","of which training:",1000))+
      # geom_smooth(method=lm)+
      xlab("Real value")+
      ylab("Computed value")+
      theme_minimal()+
      geom_abline(slope = 1,intercept = 0)

    plot(p)
    ############################
```

<br>

- The *accuracy* doesn't seem to be affected by a small number of words in the single document;

- That's a great news since it means that this package can be used to *analyze Tweets*, where a major challenge is posed by the fact that users are limited to 240 characters;
 
- So let's see a real case application of the package .

</div>



##Using ReadMe on Italian Tweets

An overview of the italian institutional crisis on the 27th of May.





<br>

<div class="columns-2">



```{r "top 20 Hashtag analysis"}
#hashtag study

#tidying dataset for lexical analysis
library(tidyverse)
library(rtweet)

library(gridExtra)

library(tidytext)
library(widyr)
library(igraph)
library(ggraph)
#--------------------------------------------------------------------------------------------Hashtag
#Tidy for hashtag study
dataset=readRDS("source/results/reduced_tweet_no_duplicati_stripped_text_separated.RDS")
#only italian language tweets
dataset=dataset%>%filter(lang=="it")





# remove punctuation, convert to lowercase.
dataset_clean <- dataset %>%
  select(hashtags) %>%
  unnest_tokens(word, hashtags)
#put in original dataframe
dataset$hashtags=dataset_clean$word

a=dataset_clean %>%
  count(word, sort = TRUE)




#plot the top 20 HASHTAGS
dataset_clean %>%
  count(word, sort = TRUE) %>%
  
  top_n(20+1) %>%
  filter(!is.na(word))%>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(x = word, y = n)) +
  geom_col(aes(alpha=n),show.legend =FALSE) +
  xlab(NULL) +
  coord_flip() +
  labs(x = "Hashtags",
       y = "Count",
       title = " Top 20 hashtags by count")+
  ggplot2::theme_minimal()

```

<br>

```{r distro parole}
#plot dei risultati
megatrain=readRDS("source/results/distro_parole_tweets.rds")


megatrain%>%
  filter(number_words<100)%>%
ggplot(aes(number_words))+
  geom_histogram(bins=50)+
  ggtitle(" words count")+
  theme_minimal()
```





## Twitter activity |Signs of polarization given by some explicit hashtags

<div align="center">
```{r "Hashtag analysis zoom"}

#hashtag study

#tidying dataset for lexical analysis
library(tidyverse)
library(rtweet)

library(gridExtra)

library(tidytext)
library(widyr)
library(igraph)
library(ggraph)
#--------------------------------------------------------------------------------------------Hashtag
#Tidy for hashtag study
dataset=readRDS("source/results/reduced_tweet_no_duplicati_stripped_text_separated.RDS")
#only italian language tweets
dataset=dataset%>%filter(lang=="it")

TIMEFRAME=data.frame(begin=as.POSIXct("2018-05-27 17:15:21", tz = "GMT"),end=as.POSIXct("2018-05-27 18:57:37", tz = "GMT"))

#zoom on hot time
p4=dataset%>%
  #filter(has_top_hashtag==TRUE)%>%
    filter(hashtags=="iostoconmattarella" | hashtags=="mattarelladimettiti" | hashtags=="impeachment" 
           | hashtags=="impeachmentmattarella" )%>%

  group_by(hashtags,is_retweet)%>%
   filter ( created_at >= as.POSIXct("2018-05-27 16:00:00",tz = "GMT") & created_at <= as.POSIXct("2018-05-28 00:00:00",tz = "GMT") ) %>%
  ts_plot( "5 minutes") +
  ggplot2::theme_minimal() +
  ggplot2::theme(plot.title = ggplot2::element_text(face = "bold")) +
  ggplot2::labs(
    x = NULL,
    y = NULL,
    title = "Frequency  Tweets statuses for period 16:00-24:00 27/05/2018",
    subtitle = "Twitter status (tweet) counts aggregated using five-minutes intervals",
    caption = "\nSource: Data collected from Twitter's REST API via rtweet",
    linetype="Retweet?",
    colour = " Hashtags"
  )  +
    geom_rect(data = TIMEFRAME,
                aes(xmin = as.POSIXct(begin,tz = "GMT"), xmax =as.POSIXct( end,tz = "GMT"), ymin = -Inf, ymax = +Inf),
                inherit.aes = FALSE, fill = "red", alpha = 0.1)
                
                
                p4
```
</div>



## Major challenge| Build the labelled trainingset {.build}

<div align="center">
Proposed solution
</div>

<br>

- *First*: Gather tweets with hashtags with explicit meaning.

<div align="center">

**#iostoconmattarella**

VS

**#mattarelladimettiti, #Impeachment, #Impeachmentmattarella**
</div>

- *Second*: Select those users that never used both of them, and retain only tweets written from them.



## Testing Readme on Italian tweets dataset | Language indipendent since we dropped stemming{.build}

<div align=center>
```{r}



mattarella1=readRDS("source/results/tweet_sentiment_test.rds")
valori_tot=mattarella1
valori_tot1=valori_tot
valori_tot1$type[valori_tot1$type==0] <- "Contra"
valori_tot1$type[valori_tot1$type==1] <- "Pro"
valori_tot1=rename(valori_tot1,Sentiment=type)
p=ggplot(data=valori_tot1,aes(x=value.real,y=value.est,label = value.est))+
  geom_point(aes(color=Sentiment))+
  #  geom_linerange(aes(ymin=value.est-value, ymax=value.est+value,color=type)) +
  # ggtitle(" Computed vs fitted values for Amazon music dataset, frasi corte")+
  ggplot2::labs(
    title = "Computed vs fitted values for Twitter dataset",
    subtitle = paste("total number of elements:",27754,",","of which training:",1000),
    caption = "Tweets downloaded usind R package Rtweet"  )+
  # geom_smooth(method=lm)+
  xlab("Real value")+
  ylab("Computed value")+
  theme_minimal()+
  geom_abline(slope = 1,intercept = 0)

plot(p)

```

</div>

<!-- -It works! -->

## Sentiment towards President Mattarella | Estimated from 10000 random sampled tweets with #Mattarella

<div align=center>

```{r Mattarella}
mattarella1=readRDS("source/results/tweet_mattarella_sentiment.rds")
valori_tot=mattarella1
valori_tot1=valori_tot
valori_tot1$type[valori_tot1$type==0] <- "Contra"
valori_tot1$type[valori_tot1$type==1] <- "Pro"

ggplot(valori_tot1,aes(y=value.est,x=type,fill=type))+
  labs(title="Pro e Contra Mattarella",
       subtitle= "From tweets with hashtag Mattarella. 10000 tweets considered, 2000 tweets as training set",
       x="Sentiment",
       y="Percentage",
       caption="Tweets gathered using R library Rtweet for the period  May 26-June 03 2018")+
  geom_col()+
  theme_minimal()+
  guides(fill=FALSE)

```

</div>


## Bibliography

[1] "A Method of Automated Nonparametric Content Analysis for Social Science", D. J. Hopkins, G. King, 2010.

[2] "Verbal Autopsy Methods with Multiple Causes of Death", King, Gary and Ying Lu, 2008.

[3] "ReadMe: Software for Automated Content Analysis", D. Hopkins, G. King, M. Knowles, S. Melendez, 2012.

[4] Amazon dataset: http://jmcauley.ucsd.edu/data/amazon/

[5] ReadMe software releases: https://gking.harvard.edu/readme

***

# Thank you for your attention
